Output Summary

### 2023-04-04 03-post-process/output-plots-children/10-noise-term-diags.Rmd
Process model qresids now in Rmd

@@

```{r pi-noise, fig.width = 6, fig.height = 5}
my_par()
tmp = lapply(1:jags_data$nj, function(j) plot_proc_qresid("Lpi_qresid[year,pop]", pop = j, main = pops[j]))
oma_labels("Brood Year", "logit(Proportion Fall Migrants) Residual")
proc_error_ests = rbind(proc_error_ests, do.call(rbind, tmp))
```

@@

```{r trib-to-LGR-surv-nor, fig.width = 6, fig.height = 5}
my_par()
tmp = lapply(1:jags_data$nj, function(j) plot_proc_qresid("Lphi_Mb_Ma_qresid[year,origin,pop]", pop = j, origin = 1, main = pops[j]))
oma_labels("Brood Year", "logit(Trib to LGR Survival [NOR]) Residual")
proc_error_ests = rbind(proc_error_ests, do.call(rbind, tmp))
```

@@

```{r trib-to-LGR-surv-hor, fig.width = 6, fig.height = 5}
my_par()
tmp = lapply(1:jags_data$nj, function(j) plot_proc_qresid("Lphi_Mb_Ma_qresid[year,origin,pop]", pop = j, origin = 2, main = pops[j]))
oma_labels("Brood Year", "logit(Trib to LGR Survival [HOR]) Residual")
proc_error_ests = rbind(proc_error_ests, do.call(rbind, tmp))
```

@@

```{r proc-error-table-prep}
# rename parameters to human-readable format
proc_error_ests$param = stringr::str_replace(proc_error_ests$param, "Lphi_E_Pb_qresid\\[year,pop\\]", "Egg to Parr Survival")
proc_error_ests$param = stringr::str_replace(proc_error_ests$param, "lL_Pb_qresid\\[year,pop\\]", "Summer Parr Length")
proc_error_ests$param = stringr::str_replace(proc_error_ests$param, "Lpi_qresid\\[year,pop\\]", "Migratory Strategy Apportionment")
proc_error_ests$param = stringr::str_replace(proc_error_ests$param, "Lphi_Pa_Mb_qresid\\[year,LH_type,pop\\]", "Overwinter Survival")
proc_error_ests$param = stringr::str_replace(proc_error_ests$param, "lDelta_L_Pb_Mb_qresid\\[year,pop\\]", "Summer to Spring 'Growth Factor'")
proc_error_ests$param = stringr::str_replace(proc_error_ests$param, "Lphi_Mb_Ma_qresid\\[year,origin,pop\\]", "Trib to LGR Survival")
proc_error_ests$param = stringr::str_replace(proc_error_ests$param, "Lphi_Ma_O0_qresid\\[year,origin\\]", "LGR to BON Survival")
proc_error_ests$param = stringr::str_replace(proc_error_ests$param, "Lphi_O0_O1_qresid\\[year,origin,pop\\]", "Yr1 Ocean Survival")
proc_error_ests$param = stringr::str_replace(proc_error_ests$param, "Lpsi_O1_qresid\\[year,origin,pop\\]", "Pr(Mature Age-3)")
proc_error_ests$param = stringr::str_replace(proc_error_ests$param, "Lpsi_O2_qresid\\[year,origin,pop\\]", "Pr(Mature Age-4)")
proc_error_ests$param = stringr::str_replace(proc_error_ests$param, "Lphi_Rb_Ra_qresid\\[year,origin\\]", "BON to LGR Survival")
proc_error_ests$param = stringr::str_replace(proc_error_ests$param, "Lphi_Sb_Sa_qresid\\[year,pop\\]", "Pre-spawn Survival")

>>>>

```{r make-process-noise-terms, results = "hide"}
# create quantile residuals for a time series from one process
# e.g., Pr(mature age-3) for all years for one population, origin, and age combination
make_qresid = function(real_param, mean_param, sig_param, f_real = qlogis, f_mean = qlogis, yrs = observable, ...) {

  # extract posterior samples of the correct quantities for all years, and apply the correct transformation where needed
  real_samps = f_real(post_subset(post, sub_index(real_param, year = yrs, ...), matrix = TRUE))  # realized value from ppd
  mean_samps = f_mean(post_subset(post, sub_index(mean_param, year = yrs, ...), matrix = TRUE))  # mean parameter ppd
  sig_samps = as.numeric(post_subset(post, sub_index(sig_param, ...), matrix = TRUE))            # sd parameter of ppd

  # remove any ^ or $ symbols used in specifying specific parameters
  real_param = postpack:::rm_regex_lock(real_param)
  mean_param = postpack:::rm_regex_lock(mean_param)

  # calculate the quantile of the realized value given the mean and sd parameter
  # performed MCMC draw by draw for each year separately
  out = sapply(1:length(yrs), function(y) {
    pnorm(
      real_samps[,sub_index(real_param, year = yrs[y], ...)],
      mean_samps[,sub_index(mean_param, year = yrs[y], ...)],
      sig_samps
    )
  })

  # create new names for the output
  # basically add "_qresid" to the end of the realized node name
  new_names = sub_index(real_param, year = yrs, ...)
  new_names = stringr::str_replace(new_names, "\\[", "_qresid[")

  # include the correct letter to indicate transformation performed (L = logit or l = log) and drop any ^
  new_names = paste0(ifelse(f_real(1) == log(1), "l", "L"), new_names)
  colnames(out) = new_names

  # return the output: a matrix of posterior samples for the desired process
  return(out)
}

# create quantile residuals for all time series from one process
# e.g., Pr(mature age-3) for all years, all populations, and age-combinations
# type = "pop" is for processes that vary by population
# type = "origin" is for processes that very by origin
make_qresid_all = function(real_param, mean_param, sig_param, f_real = qlogis, f_mean = qlogis, type = "pop", ...) {

  # print an indicator for which process is being analyzed
  process_name = postpack:::rm_regex_lock(real_param)
  process_name = postpack:::drop_index(process_name)
  cat("  ", process_name, "\n")

  # loop over populations if appropriate
  if (type == "pop") {
    qresid_out = lapply(1:jags_data$nj, function(j) {
      cat("\r    Population: ", c("CAT", "LOS", "MIN", "UGR")[j], sep = "")
      out = make_qresid(real_param, mean_param, sig_param, f_real = f_real, f_mean = f_mean, pop = j, ...)
      return(out)
    })
  }

  # loop over origins if appropriate
  if (type == "origin") {
    qresid_out = lapply(1:jags_data$no, function(o) {
      cat("\r    Origin: ", c("NOR", "HOR")[o], sep = "")
      out = make_qresid(real_param, mean_param, sig_param, f_real = f_real, f_mean = f_mean, origin = o, ...)
      return(out)
    });
  }

  # print a newline
  cat("\n")

  # combine list elements into one matrix
  qresid_out = do.call(cbind, qresid_out)

  # return the output
  return(qresid_out)
}

# container object to append output to
qresid_post = postpack:::id_mat(post)

# the years to calculate residuals for
observable = 2:jags_data$ny_obs
# "^phi_E_Pb[year,pop]""phi_E_Pb_dot2[year,pop]", "sig_Lphi_E_Pb[pop]"

# egg to parr survival
qresid_post = cbind(qresid_post,
                    make_qresid_all("^phi_E_Pb[year,pop]",
                                    "phi_E_Pb_dot2[year,pop]",
                                    "sig_Lphi_E_Pb[pop]",
                                    type = "pop")
)

# parr mean length
qresid_post = cbind(qresid_post,
                    make_qresid_all("^L_Pb[year,pop]",
                                    "^L_Pb_dot[year,pop]",
                                    "sig_lL_Pb[pop]",
                                    type = "pop", f_real = log, f_mean = log)
)

# Pr(fall migrant)
qresid_post = cbind(qresid_post,
                    make_qresid_all("^pi[year,LH_type,pop]",
                                    "mu_pi[LH_type,pop]",
                                    "sig_Lpi[pop]",
                                    type = "pop", LH_type = jags_data$i_fall)
)

# overwinter survival: fall migrants
qresid_post = cbind(qresid_post,
                    make_qresid_all("^phi_Pa_Mb[year,LH_type,pop]",
                                    "^phi_Pa_Mb_dot[year,LH_type,pop]",
                                    "sig_Lphi_Pa_Mb[pop]",
                                    type = "pop", LH_type = jags_data$i_fall)
)

# overwinter survival: spring migrants
qresid_post = cbind(qresid_post,
                    make_qresid_all("^phi_Pa_Mb[year,LH_type,pop]",
                                    "^phi_Pa_Mb_dot[year,LH_type,pop]",
                                    "sig_Lphi_Pa_Mb[pop]",
                                    type = "pop", LH_type = jags_data$i_spring)
)

# "growth" factor
qresid_post = cbind(qresid_post,
                    make_qresid_all("^Delta_L_Pb_Mb[year,pop]",
                                    "^Delta_L_Pb_Mb_dot[year,pop]",
                                    "sig_Lphi_Pa_Mb[pop]",
                                    type = "pop", f_real = log, f_mean = log)
)

# trib to LGR survival: NOR
qresid_post = cbind(qresid_post,
                    make_qresid_all("^phi_Mb_Ma[year,LH_type,origin,pop]",
                                    "^phi_Mb_Ma_dot[year,LH_type,origin,pop]",
                                    "sig_Lphi_Mb_Ma[pop]",
                                    type = "pop", LH_type = jags_data$i_spring, origin = jags_data$o_nor)
)

# trib to LGR survival: HOR
qresid_post = cbind(qresid_post,
                    make_qresid_all("^phi_Mb_Ma[year,LH_type,origin,pop]",
                                    "^mu_phi_Mb_Ma[LH_type,origin,pop]",
                                    "sig_Lphi_Mb_Ma[pop]",
                                    type = "pop", LH_type = jags_data$i_spring, origin = jags_data$o_hor)
)

# LGR to ocean survival
qresid_post = cbind(qresid_post,
                    make_qresid_all("^phi_Ma_O0[year,origin]",
                                    "mu_phi_Ma_O0[origin]",
                                    "sig_Lphi_Ma_O0[origin]",
                                    type = "origin")
)

# Yr1 ocean survival
qresid_post = cbind(qresid_post,
                    make_qresid_all("^phi_O0_O1[year,origin,pop]",
                                    "^phi_O0_O1_dot2[year,origin,pop]",
                                    "sig_Lphi_O0_O1[pop]",
                                    type = "pop", origin = jags_data$o_nor)
)

# Pr(mature age-3): NOR
qresid_post = cbind(qresid_post,
                    make_qresid_all("^psi_O1[year,origin,pop]",
                                    "^mu_psi_O1[origin,pop]",
                                    "sig_Lpsi_O1[pop]",
                                    type = "pop", origin = jags_data$o_nor)
)

# Pr(mature age-3): HOR
qresid_post = cbind(qresid_post,
                    make_qresid_all("^psi_O1[year,origin,pop]",
                                    "^mu_psi_O1[origin,pop]",
                                    "sig_Lpsi_O1[pop]",
                                    type = "pop", origin = jags_data$o_hor)
)

# Pr(mature age-3): NOR
qresid_post = cbind(qresid_post,
                    make_qresid_all("^psi_O2[year,origin,pop]",
                                    "^mu_psi_O2[origin,pop]",
                                    "sig_Lpsi_O2[pop]",
                                    type = "pop", origin = jags_data$o_nor)
)

# Pr(mature age-3): HOR
qresid_post = cbind(qresid_post,
                    make_qresid_all("^psi_O2[year,origin,pop]",
                                    "^mu_psi_O2[origin,pop]",
                                    "sig_Lpsi_O2[pop]",
                                    type = "pop", origin = jags_data$o_hor)
)

# BON to LGR survival
qresid_post = cbind(qresid_post,
                    make_qresid_all("^phi_Rb_Ra[year,origin]",
                                    "mu_phi_Rb_Ra[origin]",
                                    "sig_Lphi_Rb_Ra[origin]",
                                    type = "origin")
)

# Pre-spawn survival
qresid_post = cbind(qresid_post,
                    make_qresid_all("^phi_Sb_Sa[year,pop]",
                                    "^mu_phi_Sb_Sa[pop]",
                                    "sig_Lphi_Sb_Sa[pop]",
                                    type = "pop", origin = jags_data$o_hor)
)

qresid_post = qresid_post[,-c(1:2)]

post = post_bind(post, qresid_post)
```

@@

```{r pi-noise, fig.width = 6, fig.height = 5}
my_par()
tmp = lapply(1:jags_data$nj, function(j) plot_proc_qresid("Lpi_qresid[year,LH_type,pop]", pop = j, main = pops[j], LH_type = i_fall))
oma_labels("Brood Year", "logit(Proportion Fall Migrants) Residual")
proc_error_ests = rbind(proc_error_ests, do.call(rbind, tmp))
```

@@

```{r trib-to-LGR-surv-nor, fig.width = 6, fig.height = 5}
my_par()
tmp = lapply(1:jags_data$nj, function(j) plot_proc_qresid("Lphi_Mb_Ma_qresid[year,LH_type,origin,pop]", pop = j, origin = 1, LH_type = 2, main = pops[j]))
oma_labels("Brood Year", "logit(Trib to LGR Survival [NOR]) Residual")
proc_error_ests = rbind(proc_error_ests, do.call(rbind, tmp))
```

@@

```{r trib-to-LGR-surv-hor, fig.width = 6, fig.height = 5}
my_par()
tmp = lapply(1:jags_data$nj, function(j) plot_proc_qresid("Lphi_Mb_Ma_qresid[year,LH_type,origin,pop]", pop = j, origin = 2, LH_type = 2, main = pops[j]))
oma_labels("Brood Year", "logit(Trib to LGR Survival [HOR]) Residual")
proc_error_ests = rbind(proc_error_ests, do.call(rbind, tmp))
```

@@

```{r proc-error-table-prep}
# rename parameters to human-readable format
proc_error_ests$param = stringr::str_replace(proc_error_ests$param, "Lphi_E_Pb_qresid\\[year,pop\\]", "Egg to Parr Survival")
proc_error_ests$param = stringr::str_replace(proc_error_ests$param, "lL_Pb_qresid\\[year,pop\\]", "Summer Parr Length")
proc_error_ests$param = stringr::str_replace(proc_error_ests$param, "Lpi_qresid\\[year,LH_type,pop\\]", "Migratory Strategy Apportionment")
proc_error_ests$param = stringr::str_replace(proc_error_ests$param, "Lphi_Pa_Mb_qresid\\[year,LH_type,pop\\]", "Overwinter Survival")
proc_error_ests$param = stringr::str_replace(proc_error_ests$param, "lDelta_L_Pb_Mb_qresid\\[year,pop\\]", "Summer to Spring 'Growth Factor'")
proc_error_ests$param = stringr::str_replace(proc_error_ests$param, "Lphi_Mb_Ma_qresid\\[year,LH_type,origin,pop\\]", "Trib to LGR Survival")
proc_error_ests$param = stringr::str_replace(proc_error_ests$param, "Lphi_Ma_O0_qresid\\[year,origin\\]", "LGR to BON Survival")
proc_error_ests$param = stringr::str_replace(proc_error_ests$param, "Lphi_O0_O1_qresid\\[year,origin,pop\\]", "Yr1 Ocean Survival")
proc_error_ests$param = stringr::str_replace(proc_error_ests$param, "Lpsi_O1_qresid\\[year,origin,pop\\]", "Pr(Mature Age-3)")
proc_error_ests$param = stringr::str_replace(proc_error_ests$param, "Lpsi_O2_qresid\\[year,origin,pop\\]", "Pr(Mature Age-4)")
proc_error_ests$param = stringr::str_replace(proc_error_ests$param, "Lphi_Rb_Ra_qresid\\[year,origin\\]", "BON to LGR Survival")
proc_error_ests$param = stringr::str_replace(proc_error_ests$param, "Lphi_Sb_Sa_qresid\\[year,pop\\]", "Pre-spawn Survival")
